# Bill of Materials

```{r}
#| results: "asis"
#| echo: false
source("_common.R")
library(janitor)
library(readxl)
```

Each installation has its own BOM structure report. 

Column definitions:

- **Lv**: The hardware level, which can expand to multiple monitoring systems if needed.
- **Component**: Serial number to be scanned if needed.
- **Seq**: The operation sequence from installation onward.
- **Description**: Description of the part or component.
- **Quantity per**: Number of parts needed for functionality.
- **Fix**: Repairability of parts.
- **On-date**: Used to track stable parts. Default to "Prepared" date on top.
- **Off-date**: Used to track depreciated parts. NULL if stable.
- **Rev**: Revision to track prints and schematics.
- **Unit Cost**: Cost at the time of document.
- **Vendor**: Purchase location.
- **Vendor PN**: Vendor part number.


```{r}
file_list <- list.files("../BOM/", pattern = "BOM STRUCTURE REPORT – ALL LEVELS", full.names = TRUE)

bom_data <- data.frame()

for (filename in file_list) {
  location_id <- stringr::str_extract(filename, "[A-Z]\\d{2}\\.\\d{2}(?=\\.xlsx$)")
  
  data <- read_xlsx(filename, skip = 6) |>
    select(-matches("^\\.\\.")) |> 
    mutate(location_id = location_id) 
  
  bom_data <- bind_rows(bom_data, data)
}

bom_data <- bom_data |> relocate(location_id) 

bom_data |> filter(location_id == "A00.01")

```

The total cost per installation: 

```{r}
bom_data <- bom_data |> 
  clean_names() |> 
  mutate(unit_cost = parse_number(unit_cost))

bom_data |>
  group_by(location_id) |>
  summarize(total_cost = sum(unit_cost, na.rm = T))

```

We can also plot the parts installed. 

```{r}
bom_data |> 
  group_by(location_id, component, description) |> 
  count() |> 
  ggplot(aes(component, n, fill = location_id)) +
  geom_col() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
```

If there are any one offs, you could filter through the original table using the serial number to track parts. 



