# Query Data 

```{r}
#| results: "asis"
#| echo: false
source("_common.R")

library(DBI)
library(RPostgres)

```

Data is stored in two AWS RDS databases. Please let me know if you have lost the credentials.

1. gd_phm_status : This has the model output information, hardware maintenance information, versioning, and configuration settings. 

2. metrics : This hosts accuracy and reliability metrics for all my projects. The table for this project is called "gd_phm".


Shown below is a generic sequence of how one might pull data.

## Make a Connection

```{r}
con <- dbConnect(
  RPostgres::Postgres(),
  dbname = "gd_phm_status",
  host = Sys.getenv("gd_db_host"),
  port = 5432,
  user = Sys.getenv("gd_db_user"),
  password = Sys.getenv("gd_db_pw")
)
```

## Get Relevant Table Names

```{r}
tbl <- DBI::dbGetQuery(
  conn = con,
  "SELECT table_name
  FROM information_schema.tables
  WHERE table_name like '%gd%'
"
)

tbl |> pull(table_name) |> cat(sep = ", ")
```

## Pull Data

```{r}
tbl <- DBI::dbGetQuery(
  conn = con,
  
  "WITH joined_tbl AS (
    SELECT
      gd_config_hw_tbl.*,
      meas_date,
      meas_seq,
      status,
      gd_meas_tbl.model AS model,
      prob,
      model_version,
      version_from,
      version_to
    FROM gd_config_hw_tbl
    INNER JOIN gd_meas_tbl
      ON (gd_config_hw_tbl.location_id = gd_meas_tbl.location_id)
    INNER JOIN gd_config_models_tbl
      ON (
        gd_config_hw_tbl.location_id = gd_config_models_tbl.location_id AND
        gd_meas_tbl.model = gd_config_models_tbl.model
         )
    )
  
  SELECT
    state,
    city_code,
    location_id,
    model,
    model_version,
    meas_date,
    MAX(prob) AS prob_max
  FROM joined_tbl
  WHERE (meas_date >= version_from AND ((version_to IS NULL) OR meas_date <= version_to))
  GROUP BY state, city_code, location_id, model, model_version, meas_date;"
)

tbl |> head(10)

```


## Plot Data

```{r}
tbl |> 
  filter(meas_date < as.Date("2023-06-11")) |> 
  mutate(model = ifelse(model == "arm_brkt", "Arm Bracket Fault", model)) |> 
  ggplot(aes(meas_date, prob_max, col = model_version)) +
  geom_point() +
  facet_grid(location_id~ model, scales = "free") +
  scale_x_date(date_breaks = "2 weeks") +
  labs(
    title = "Daily Maximum Fault Probability for Each IoT Device Monitored Against Model Version", 
    x = "Date", 
    y = "Fault Probability",
    col = "Model Version"
      )
```

## Important! 

Please close the connection after you are done. 

```{r}
dbDisconnect(con)
```

