# Preprocessing Methods
```{r}
#| results: "asis"
#| echo: false
source("_common.R")

```

## Example Signal

Shown below is a signal for a complete cycle. Note that the acceleration of the y axis (pointing down), is not at 1g. This is an acceptable condition. 

```{r}
d <- pre_proc_tbl("2023-06-10_garagedoor-1") 

d |> 
  plot_ad() +
  labs(title = "Accelerometer Signal (Triaxial)")

```


## Median Filter 

![](./man/figures/lifecycle-maturing.svg)

The data has irregular spike patterns caused by the slack of the curved door arm. A simple filter is used to remove them.

### Verify Filter

The choice of n should be the minimum odd value that will remove added spikes from the baseline signal.

```{r}
set.seed(123)
t <- seq(0, 10, length.out = 1000)

imp_sig <- rpois(length(t), lambda = 0.1) 
imp_sig_damp <- imp_sig * 0.01
signal <- exp(0.2 * -1*exp(-t)) + imp_sig_damp

mmed <- function(x,n=9){runmed(x,n)} 

signal_fil <- mmed(signal, n = 7) 

d_med <- tibble(
  Original = signal,
  Filtered = signal_fil
) %>%
  mutate(t = seq_along(Original)) %>%
  pivot_longer(cols = c(Original, Filtered)) |> 
  mutate(name = fct_relevel(name, c("Original", "Filtered")))

d_med |> ggplot(aes(t, value)) +
  geom_point() +
  facet_wrap(~name) +
  labs(title = "Median Filter Test on a Gompertz Curve")

```

### Filter Data

The choice of n should be the minimum odd value that will produce the final segmentation on the baseline signal. The baseline signal is the calibrated, initial measurement. This parameter needs to be verified during the qualification process.

```{r}
d <- d |> mutate(across(c(x, y, z), ~mmed(x = ., n = 5)))

d |> 
  plot_ad() +
  labs(title = "Filtered Signal")
```
Note that this process removes some information, and it is not the signal that is fed to the models.

## Grenerate Virtual Recipe Steps

![](./man/figures/lifecycle-maturing.svg)

The following is an automated procedure that cleans and segments the signal into idle, turn, and impulse responses. 

### Extract Turn Response

![](./man/figures/lifecycle-stable.svg)

The first step is to smooth the signal, and a machine learning method is used to isolate the accelerometer response around a turn. 


```{r}
d <- d |> 
  mutate(z_smo = loess(z ~ time, span = 0.1)$fitted, 
         y_smo = loess(y ~ time, span = 0.1)$fitted) 

ts_data <- d |> pull(z_smo) |> ts()

d |> 
  pivot_longer(c(z, z_smo, y, y_smo)) |> 
  ggplot(aes(time, value)) +
  geom_point(size = 0.5) +
  facet_wrap(~name) +
  labs(title = "Smoothed Signal")
```

```{python}
import statsmodels.api as sm1
import numpy as np
import pandas as pd
import matplotlib.pylab as plt
from sklearn import linear_model, datasets

# Return the t-statistic for a given parameter estimate.
def tValLinR(close):
    # tValue from a linear trend
    x = np.ones((close.shape[0], 2))
    x[:, 1] = np.arange(close.shape[0])
    ols = sm1.OLS(close, x).fit()
    return ols.tvalues[1]


def getBinsFromTrend(molecule, close, span):
    '''
    Derive labels from the sign of t-value of trend line
    output includes:
      - t1: End time for the identified trend
      - tVal: t-value associated with the estimated trend coefficient
      - bin: Sign of the trend
    The t-statistics for each tick has a different look-back window.

    - idx start time in look-forward window
    - dt1 stop time in look-forward window
    - df1 is the look-forward window
    - iloc ?
    '''
    out = pd.DataFrame(index=molecule, columns=['t1', 'tVal', 'bin', 'windowSize'])
    hrzns = range(*span)
    windowSize = span[1] - span[0]
    maxWindow = span[1] - 1
    minWindow = span[0]
    for idx in close.index:
        idx += maxWindow
        if idx >= len(close):
            break
        df_tval = pd.Series(dtype='float64')
        iloc0 = close.index.get_loc(idx)
        # if iloc0+max(hrzns) > close.shape[0]:
        #    continue
        for hrzn in hrzns:
            dt1 = close.index[iloc0 - hrzn + 1]
            df1 = close.loc[dt1:idx]
            df_tval.loc[dt1] = tValLinR(df1.values)  # calculates t-statistics on period
            dt1 = df_tval.replace([-np.inf, np.inf, np.nan],
                              0).abs().idxmax()  # get largest t-statistics calculated over span period

        # print(df_tval.index[-1])
        # print(dt1)
        # print(abs(df_tval.values).argmax() + minWindow)
        out.loc[idx, ['t1', 'tVal', 'bin', 'windowSize']] = df_tval.index[-1], df_tval[dt1], np.sign(df_tval[dt1]), abs(
            df_tval.values).argmax() + minWindow  # prevent leakage
    out['t1'] = pd.to_datetime(out['t1'])
    out['bin'] = pd.to_numeric(out['bin'], downcast='signed')

    # deal with massive t-Value outliers - they dont provide more confidence and they ruin the scatter plot
    tValueVariance = out['tVal'].values.var()
    tMax = 20
    if tValueVariance < tMax:
        tMax = tValueVariance

    out.loc[out['tVal'] > tMax, 'tVal'] = tMax  # cutoff tValues > 20
    out.loc[out['tVal'] < (-1) * tMax, 'tVal'] = (-1) * tMax  # cutoff tValues < -20
    return out.dropna(subset=['bin'])



idx_range_from = 5
idx_range_to = 10
df0 = pd.Series(r.ts_data)

span = [idx_range_from,idx_range_to,1] # [3,10,1] = range(3,10)

df1 = getBinsFromTrend(df0.index, df0, span) 
tValues = df1['tVal'].values
```

```{r}
library(reticulate)

d_f <- py$df1 |> 
  rownames_to_column("id") |> 
  unnest(tVal) 

d_fmerg <- d |> 
  rownames_to_column("id") |> 
  left_join(d_f) |>
  tidyr::fill(bin, .direction = "downup") |> 
  mutate(id = as.numeric(id)) |> 
  na.omit() |> 
  mutate(run_id = consecutive_id(bin)) |> 
  group_by(run_id) |> 
  mutate(bin = ifelse(n() < 1000, 0, bin))

d_fmerg |> 
  ggplot(aes(id, z, col = as.factor(bin))) +
  geom_point(size = 0.3) +
  labs(x = "time (s)", 
       y = "Acceleration (g)",
       col = "Recipe Step")

```

Legend definitions:

- 0: Horizontal or vertical position
- -1: Turn, coming down
- 1: Turn, going up


### Extract Door Idle Response

![](./man/figures/lifecycle-questioning.svg)

The following is a dependent mixture model to detect when the door is idle.

```{r}
library(depmixS4)

set.seed(123)

hmm_model <- depmixS4::depmix(data = d_fmerg, nstates = 2, y_smo~y)
hmm_fit <- fit(hmm_model)

# hmm_fit@transition

# plot(ts(posterior(hmm_fit, type = "smoothing")), ylab = "probability", frame = FALSE)
```


## Final Results

```{r}
d_fmerg_final <- d_fmerg |> 
  ungroup() |> 
  bind_cols(smo_prob = posterior(hmm_fit, type = "smoothing")[,1]) |> 
  mutate(bin_hmm = ifelse(smo_prob > 0.999, "Transition", "Idle")) |> 
  mutate(run_idle_id = consecutive_id(bin_hmm)) |> 
  group_by(run_idle_id) |> 
  mutate(numbers = n(),
         bin_hmm = ifelse(n() < 2000 , "Transition", bin_hmm)) |> 
  mutate(recipe_step = paste(bin_hmm, bin))
```

```{r}
d_fmerg_final |> 
  pivot_longer(c(x, y, z)) |>
  ggplot(aes(time, value, col = as.factor(recipe_step))) +
  geom_point(size = 0.3) +
  facet_wrap(~name) +
  labs(x = "time (s)",
       y = "Acceleration (g)",
       col = "Recipe Step")
```

The sensor position around the turn (pink, door going up) has a distinct signature compared to the opposite (green, door coming down). 

The shape of the signal comes from the rate of tilt around the turn, denoted by \(\theta(t)\), and can be expressed by a the following first order differential equation:

$$
\frac{d\theta}{dt} = f(\theta, G, t),
$$

where \(f(\theta, G, t)\) represents the function describing the dynamics of the garage door, and G is the geometry of the system. This function is unknown, but can be expressed for the simplest case as

$$
 \frac{d\theta}{dt} =\dot{\theta}\ = k\theta, 
$$

which states that the rate of change of the angular position of the accelerometer Y and Z axis is proportional to the current angular position ($\dot{\theta}/θ = k$). The general solution to this equation is

$$
\theta \ =  Ce^{kt},
$$
where C is a constant that depends on the initial conditions of the system. This is one of the reasons why the adjustment for the angle sweep using the adjustable mount is important.



